Self-Supervised Transfer Learning From 2D Images to 3D Point Clouds

### 2022-05-09 doc/paper.tex
adds some analysis sections

We first test the case of pre-training our model on ScannetV2 and fine-tuning on S3DIS. Detailed results are shown in table \ref{table:s3disResults}. Fully supervised pre-training has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance. Almost all base algorithms improve downstream results, except DepthContrast. Our algorithm is comparable to PointContrast, both achieving a 1\% mIOU increase but perform worse than CSC.

@@

However, looking at the effects on Scannet in table \ref{table:scannetVaryingDataAmount}, all methods struggle to improve upon training from scratch. In many cases they even perform substantially worse, with out method not being able to improve in almost all scenarios. Contrastive Scene Contexts is the only method that shows some consistent improvement. It is interesting to note that even though semantic segmentation receives a boost, this does not necessarily translate into instance segmentation.

@@

@@

Talk about graphs showing that you can get close to final performance within a fraction of the time.

@@

@@

\begin{itemize}
    \item that transferring 2D features to 3D improves performance across different domains.
    \item that although some algorithms perform extremely well in some domains, they fail in others. Or they might work well on one task but not another.
    \item PointContrast \& CSC both ideally require overlapping scenes to perform properly whereas our algorithm does not.
    \item Pre-training does little when large labeled datasets are used but are more effective when datasets are very small
    \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
    \item Overal can expect about a 1\% increase in performance
    \item Our method, which is easy to implement and does not require a pose transformation between frames, is a good start in trying to find best performance.
    \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
    \item It may be worth trying a self-supervised method to see performance boost before spending extra time labeling the entire dataset
    \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
\end{itemize}

>>>>

We first test the case of pre-training our model on ScannetV2 and fine-tuning on S3DIS. Detailed results are shown in table \ref{table:s3disResults}. Fully supervised pre-training on Scannet has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance. Almost all base algorithms improve downstream results, except DepthContrast. Our algorithm is comparable to PointContrast, both achieving a 1\% mIOU increase but perform worse than CSC.

@@

However, looking at the effects on Scannet in table \ref{table:scannetVaryingDataAmount}, all methods struggle to improve upon training from scratch. In many cases they even perform substantially worse, with our method not being able to improve in almost all scenarios. Contrastive Scene Contexts is the only method that shows some consistent improvement. It is interesting to note that even though semantic segmentation receives a boost, this does not necessarily translate into instance segmentation.

@@

Since our method does not rely on registered point clouds, we wanted to see how sensitive PointContrast and CSC are to using the same scan for both embeddings. We present out findings in tables \ref{table:s3disResults} and \ref{table:scannetVaryingDataAmount}. We found that for S3DIS using the same scan only performed worse overall however was able to do better in a few areas. Contrastive Scene Contexts also saw a small drop in performance. This was also noted in \cite{xie2020pointcontrast}. However, using a smaller minimum overlap ratio in most cases improved performance. Our method is easier to implement as it does not rely on knowing the transformation between point-clouds and generally gets more consistent increases across the range on available data labels than using the same scan for PointContrast and Contrastive Scene contexts.

\subsection{Pre-training on Out-of-Distribution data}
\label{sec:results:varying}

Although the performance of using Scannet as the pre-training and downstream dataset has relatively low gains, those same weights have a much larger impact on S3DIS and consistently improve on KITTI across all data ratios. This discrepancy suggests that using a pre-training dataset which is out-of-distribution of the downstream dataset has a more consistent and larger performance gain than using the same dataset. This is potentially because using the same dataset already exposes the network to the same information, thus it is not learning as much new information as it could from a different dataset.

Also, on KITTI, when using more of the data we see a larger performance improvement. Since the pre-training is done on different data, the algorithms benefit more from a larger labeled dataset. Unlike when using the same dataset for both pre-training and training where the two overlap more.

@@

Table \ref{fig:validationComparison} shows the validation curves over the training steps of training from scratch vs other methods. All methods show an immediate improvement early on in the training cycle. This is potentially useful during development on downstream tasks as the final performance can be known within the first 2000 cycles instead of 20000. In S3DIS for example, algorithms like point contrast can get to within 1 mIOU of the final scratch performance in just 1000 cycles. This can drastically reduce training time and development time.

@@

Across all of our experiments, its been impossible to see a consistent improvement in performance of each method across different datasets and labeled data ratios. For example, DepthContrast might get the largest increase by far when using 5\% of semantic Scannet labels but fails when applied to Scannet.

@@

\begin{itemize}
    \item that transferring 2D features to 3D improves performance across different domains.
    \item that although some algorithms perform extremely well in some domains, they fail in others. Or they might work well on one task but not another.
    \item PointContrast \& CSC both ideally require overlapping scenes to perform properly whereas our algorithm does not.
    \item Pre-training does little when large labeled datasets are used but are more effective when datasets are very small
    \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
    \item Overal can expect about a 1\% increase in performance
    \item Our method, which is easy to implement and does not require a pose transformation between frames, is a good start in trying to find best performance.
    \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
    \item It may be worth trying a self-supervised method to see performance boost before spending extra time labeling the entire dataset
    \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
    \item Weights that work on one task could be detrimental to performance on another task.
\end{itemize}
